---
title: "ANZ_Module_2_Predictive Analytics"
author: "Chanyanart KiattipornOpas"
date: "11/21/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
    toc: yes
    toc_float: yes
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# About the Task 
This task focus on **Predictive Analysis** aim to predict **The Annual Salary** of each customers. 

Explore correlations between annual salary and various customer attributes (e.g. age). These attributes could be those that are readily available in the data (e.g. age) or those that you construct or derive yourself (e.g. those relating to purchasing behaviour). Visualise any interesting correlations using a scatter plot.

Build a simple regression model to predict the annual salary for each customer using the attributes you identified above

How accurate is your model? 

For a challenge: build a decision-tree based model to predict salary. Does it perform better? How would you accurately test the performance of this model?

# Preparation 

## 1. Download Library 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(corrplot)
library(rpart)
```

## 2. Import Data into Rstudio 

```{r message=FALSE, warning=FALSE}
transac <- read_csv("ANZ_clean_data.csv")
```

## 3. Planning for Analysis  

Predict "Annual Salary" (Y) by using these attributes: 

    - Age
    - Pay Frequency 
    - Frequency of Sales Transaction
    - Mean of Sales Transaction across 3 months 
    - Highest price of Sales Transaction 
    - Median Balance
    
```{r}
# Create the New List for each Customers by id 

customers <- transac %>% 
                select(customer_id, age) %>%
                group_by(customer_id) %>% 
                distinct() %>%
                as.data.frame()

head(customers)
```

# Derive the attributes from Data 

## 1. Salary 

```{r}
# Extract only PAY/SALARY transaction type. 

pay_salary <- transac %>% 
    select(customer_id, 
           txn_description, 
           date,
           amount) %>%
    filter(txn_description == "PAY/SALARY") %>% 
    group_by(customer_id) %>%
    as.data.frame()

head(pay_salary)
```

## 2. Pay Frequency 

```{r}
pay_salary$date <- as.Date(pay_salary$date) 

# How many time of each customers got paid in this 3 months? 

pay_freq <- pay_salary %>% 
    group_by(customer_id) %>% 
    count(amount) %>%
    as.data.frame()

# rename columns
colnames(pay_freq)[3] <- "pay_frequent"

tail(pay_freq)
```

## 3. Annual Salary (Y)
```{r}
# Total Salary across 3 months 

total_3m_salary <- pay_freq %>% 
    mutate(threeM_salary = amount*pay_frequent)

head(total_3m_salary)
```

Three months is one Quater, so 1 year = 4 quaters
```{r}
# Calculate Annual Salary 

annual_salary <- total_3m_salary %>% 
    mutate(annual_salary = threeM_salary*4)

# rename columns
colnames(annual_salary)[2] <- "salary"

tail(annual_salary)
```


```{r}
customers <- customers %>% 
    inner_join(annual_salary, by = "customer_id")

tail(customers)
```

## 4. Frequency / Mean / Highest price of Sales Transaction across 3 months 

```{r}
sale_transac <- transac %>% select(customer_id, 
                            txn_description, 
                            amount) %>%
            filter(txn_description == "POS" |
                     txn_description == "SALES-POS") %>%
            group_by(customer_id) %>%
            summarise(transac_freq = length(customer_id),
                      sum_amount = sum(amount),
                      mean_amount = mean(amount),
                      highest_amount= max(amount))

head(sale_transac)           
```

```{r}
customers <- customers %>% 
    inner_join(sale_transac, by = "customer_id") 

head(customers)
```

## 5. Balance 

```{r}
med_balance <- transac %>% select(customer_id, 
                            balance) %>% 
                  group_by(customer_id) %>%
            summarise(median_balance = median(balance)) 
```


```{r}
customers <- customers %>% 
    inner_join(med_balance, by = "customer_id") %>%
    as_tibble()

head(customers)
```

# Correlation 

```{r}
cor_matrix <- customers %>% 
                select(annual_salary, 
                       age, 
                       pay_frequent, 
                       transac_freq, 
                       mean_amount, 
                       highest_amount,
                       median_balance) %>% 
                cor() 
```

```{r}
# Visualize

corrplot(cor_matrix, 
         method = "color",
         type="upper",
         addCoef.col = "black", number.cex = 0.5, 
         order = "hclust")

```

# Check Complete Observation 

```{r}
mean(complete.cases(customers))
```
1 means all observations has no NULL value. All data are completed. 

# Prediction Model 

```{r}
# Split data 

set.seed(21) 
n <- nrow(customers)
id <- sample(1:n, size = n*0.8) 
train_data <- customers[id, ]
test_data <- customers[-id, ]

```

## 1. Linear Regression Model 

### 1.1 Train Model 
```{r}
lmModel <- lm(annual_salary ~ age + 
                              pay_frequent +
                              transac_freq +
                              mean_amount +
                              highest_amount +
                              median_balance, 
              data = train_data)

lmModel
```

### 1.2 Score Model (Prediction) 

```{r}
p1 <- predict(lmModel, newdata = test_data)
p1
```

### 1.3 Evaluate Model 

```{r}
error1 <- p1 - test_data$annual_salary
error1
```

```{r}
# RMSE 

rmse_test1 <- sqrt(mean(error1**2))
rmse_test1
```

## 2. Decision Tree Model 

### 2.1 Train Model 
```{r}
rpartModel <- rpart(annual_salary ~ age + 
                              pay_frequent +
                              transac_freq +
                              mean_amount +
                              highest_amount +
                              median_balance, 
              data = train_data)

rpartModel
```

### 2.2 Score Model (Prediction) 

```{r}
p2 <- predict(rpartModel, newdata = test_data)
p2
```

### 2.3 Evaluate Model 

```{r}
error2 <- p2 - test_data$annual_salary
error2
```

```{r}
# RMSE 

rmse_test2 <- sqrt(mean(error2**2))
rmse_test2
```

```{r}
## Compare Prediction Model with RMSE 
cat("RMSE of Linear Regression Model: ", rmse_test1,
    "\nRMSE of Decision Tree Model: ", rmse_test2)
```

# Conclusion 

To predict **The Annual Salary** of each customers. We use 6 attributions to predict the annual salary, which are customer's age, the frequent of salary pay, the median balance of customer's bank account, the purchasing behaviors(the frequent of buying, the highest and the average amounts of transactions) in this three months. 

With 2 Prediction Models that are "Linear Regression Model" and "Decision Tree Model". As the result, we found that Decision Tree Model is better than Linear Regression Model for this data. We clearly see that RMSE of Decision Tree Model (RMSE: 20169.15) less than Linear Regression Model (RMSE: 29121.63). 

However RMSE is more than 20000, that indicates an inaccuracy of the model. The variable may not suit to predict the the Annual Salary, More data and More variable are required to develop the reliable model.